---
title: "10X Example"
site: workflowr::wflow_site
output:
  workflowr::wflow_html:
    theme: yeti
    highlight: tango
    toc: true
    toc_float: 
      collapsed: true
      smooth_scroll: true
    fig_width: 6
    fig_height: 4
---

## Load Package

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      comment=NA,
                      warning = FALSE,
                      message = FALSE)
library(MultiAssayExperiment)
library(ggplot2)
library(plyr)
library(irlba)
library(matrixStats)
library(inline)
library(quadprog)
library(dplyr)
library(reshape)
library(caret)
library(fossil)
library(pracma)
library(scatterplot3d)
library(igraph)
library(Rtsne)
library(diceR)
library(gplots)
library(Matrix)
library(plotly)
library(SCNoisyClustering)
library(SingleCellExperiment)
library(SC3)
library(gridExtra)
```

## Read Data : PBMC 27K

Read data and keep the gene names separately.

```{r read_data}
orig = readMM('data/unnecessary_in_building/pbmc3k/matrix.mtx')
orig_genenames = read.table('data/unnecessary_in_building/pbmc3k/genes.tsv',
                            stringsAsFactors = FALSE)$V2
```


## Quality Control and Cell Selections

First, "explore.data" function shows the distribution of the number of UNI, detection rate, and the proportion of mitochondrial genes for each cell. Then users can decide on the cut-offs to filter the cells using "cell.filter" function, which returns the indices of cells that satisfy the input criteria. 

```{r cell_filter}
summary = SCNoisyClustering::explore.data(orig, orig_genenames)
ind = cell.filter(summary, 
                  nUMI.thresh = c(0,10000), 
                  det.rate.thresh = c(0.02, 0.067), 
                  percent.mito.thresh = c(0,0.1))
tmpX = orig[,ind]
```



## Gene Filtering

Next find variable genes using normalized dispersion. First, "plot.dispersion" function shows the distribution log of normalized dispersion (variance to mean ratio) against the gene means. The convention is to normalize the dispersion by converting them to z-scores, but one can also use median and MAD (median absolute deviation) which is less sensitive to outliers. Based on the scatterplot, users can decide on the cut-offs for the mean expression value and the dispersion to use "gene.filter" function. "gene.filter" function returns the new expression level matrix of filtered genes and the corresponding gene names. 

```{r basic_gene_filtering}
disp = plot.dispersion(X = tmpX, 
                       genenames = orig_genenames, 
                       bins=NA, 
                       median = FALSE, 
                       outliers.mean.thresh = c(30,Inf),
                       outliers.vmr.thresh = c(3,Inf))
X = gene.filter(tmpX, orig_genenames, disp, 
                mean.thresh=c(0.001, Inf), 
                dispersion.thresh = c(0.5, Inf))
genenames = X$genenames
X = X$X
```


## UMI Normalization

Use quantile-normalization to make the distribution of each cell the same. "quantile.normalize" function also performs the log transformation. 

```{r normalize_library}
X = quantile.normalize(as.matrix(X))
```

## Run SLSL

Run the clustering algorithm SLSL based on the filtered matrix. There are three main parameters. First, "klist" determines the number of neighbors to use to build initial similarity matrix, and around 1/10 of the number of cells is appropriate. "sigmalist" determines the nonlinear structure of Gaussian kernel, and the default of 1, 1.5, and 2.5 work well. The kernel_type determines the distance measure. The four options are "spearman", "pearson", "euclidean", and "combined" which combines information across three different distance measures. 

```{r slsl, cahce=TRUE}
out = SLSL(X,
           kernel_type = "combined",
           verbose=FALSE)
plot(out$tsne$Y, col=rainbow(length(unique(out$result)))[out$result])
```

<!-- ## Known Markers -->

<!-- If there are known markers, you can check how differentially they are expressed in the clustering result. The markers used below are from the supplementary table of (Schelker et al., 2017) -->

```{r biomarkers, eval = FALSE, echo = FALSE}
sorted_type = sort(out$result, index.return=TRUE)$ix

markers = c('CD3D','CD3E','CD3G','CD27','CD28',
            'CD4',
            'CD8B',
            'CD4','FOXP3','IL2RA','CTLA4',
            'CD19','MS4A1','CD79A','CD79B','BLNK',
            'CD14','CD68','CD163','CSF1R','FCGR3A',
            'IL3RA','CLEC4C','NRP1',
            'FCGR3A','FCGR3B','NCAM1','KLRB1','KLRB1','KLRC1','KLRD1','KLRF1','KLRK1',
            'VWF','CDH5','SELE',
            'FAP','THY1','COL1A1','COL3A1',
            'WFDC2','EPCAM','MCAM',
            'PMEL','MLANA','TYR','MITF'
)
ind = which(orig_genenames$V2 %in% markers)
par(mfrow = c(2,3))
for (i in 1:length(ind)){
  if(sum(tmpX[ind[i], sorted_type]!=0) > 50){
      plot(tmpX[ind[i], sorted_type], col = rainbow(11)[as.factor(out$result[sorted_type])],
       main = orig_genenames[ind[i],2], cex=0.5,
       xlab = 'cells', ylab = 'pre-processed log expression')
  }
}
```

<!-- Based on the biomarkers, we can infer that B cells are red, T cells are orange, NK cells are yelow, and monocytes are green.  -->

## Differentially Expressed Genes

Using Kruskal test, we order the p-values to find the top differentially expressed genes. Below we present 6. 

```{r DE}
degenes = de.genes(X, genenames, out$result, top.n=100, plot=12)
head(degenes)
```


## Markers for Each Cluster

```{r ggplots2, echo=FALSE}
bio.markers = find.markers(X, genenames, out$result, out$tsne$Y, top.n = 50, plot.n=3)
bio.markers$plots[[1]]
head(bio.markers$markers[[1]])

bio.markers$plots[[2]]
head(bio.markers$markers[[2]])

bio.markers$plots[[3]]
head(bio.markers$markers[[3]])
```

